---
title: "Homework 2"
author: "Nick Kurtansky"
date: "1/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Inverse CDF method for random number generation
Weibull $(a, b)$ distribution, parameters $a > 0$ and $b > 0$ has the following probability density function:

$$f(x)=ab^{-a}x^{a-1} e^{-(x/b)^a},~x>0.$$

#### 1.a Obtain the cumulative density function (CDF).

$F(x) = 1 - e^{-(x/b)^a}$

for x > 0, 0 elsewhere.

#### 1.b Solve $U = F(X)$ for $X$.

$e^{-(x/b)^a} = 1-U$

$(x/b)^a = ln(1/(1-U))$

$x = b *\sqrt[a]{ln(1/(1-U))}$n

#### 1.c Suppose that $U=0.7$ is generated. Given $a=1$ and $b=1$, What is $X$?

$x= 1 (ln(1/(1-.7)) = ln(1/.3) = 1.20397$

Note: Since $1 – U \sim U(0, 1)$ as well, can replace $1 – U$ by $U$ to get the final algorithm.

#### 1.d Draw the histogram from 1000 sampled Weibull $(a=1,b=1)$ random variables based on 1.b.
```{r}
u.sample <- runif(1000,0,1)
x.sample <- c()
for(u in u.sample){
  x <- (log(1/u))
  x.sample <- c(x.sample, x)
}
hist(x.sample, nclass=40, main="Estimated PDF of Weibull(a=1, b=1), n=1000")
```


## Problem 2: Inverse CDF method for random number generation

x  | p(x) | F(x) | U
---- | ---- | ---- | ----
-1 | .6 | .6 | [0,.6)        
2.5 | .3 | .9 | [.6, .9)
4 | .1 | 1 | (.9,1]

If $U=0.63$, then $X = 2.5$.


## Problem 3

Determine the number of days between the following datas: January 1 in the year 2018, and January 1 in the year 2028.

#### 3.a Write your own R function called, countdays.
```{r}
countdays <- function(date1,date2){
  # number of days in each month
  month.days <- c(31,28,31,30,31,30,31,31,30,31,30,31)
  
  # Break apart the dates into days/month/years components
  date1.d <- as.numeric(substr(date1, 1, 2))
  date1.m <- as.numeric(substr(date1, 4, 5))
  date1.y <- as.numeric(substr(date1, 7, 10))
  date2.d <- as.numeric(substr(date2, 1, 2))
  date2.m <- as.numeric(substr(date2, 4, 5))
  date2.y <- as.numeric(substr(date2, 7, 10))
  
  # Initialize the number of days
  num.days <- 0
  
  # Find the number of full years between them
  year.dif <- date2.y - date1.y
  num.days <- num.days + 365*year.dif

  # Find the difference in the month component
  month.dif = date2.m - date1.m
  days.of.month.dif <- sum(month.days[1:date2.m])-sum(month.days[1:date1.m])
  num.days <- num.days + days.of.month.dif

  # Find the difference in the day component
  day.dif = date2.d - date1.d
  num.days <- num.days + day.dif
  
  # Account for leap years
  # Find eligible years
  years.covered <- seq(from=date1.y, to=date2.y, by=1)
  if(date1.m > 2){
    years.covered <- years.covered[-1]
  }
  if(date2.m < 2 | (date2.m == 2 & date2.d < 29)){
    years.covered <- years.covered[-length(years.covered)]
  }
  # Find number of leap years in remaining interval
  num.leaps <- 0
  for(y in years.covered){
    if(y %% 4 != 0){
      next
    }
    if(y %% 100 != 0){
      num.leaps <- num.leaps + 1
    } else if(y %% 400 == 0){
      num.leaps <- num.leaps + 1
    }
  }
  num.days <- num.days + num.leaps
  return(num.days)
}

date.1 <- "01-01-2018"
date.2 <- "01-01-2028"
countdays(date.1, date.2)
```

#### 3.b Use the built-in R function as.Date to compute.
```{r warning=FALSE}
date.1 <- "01-01-18"
date.2 <- "01-01-28"
as.Date(date.2,"%m-%d-%y") - as.Date(date.1,"%m-%d-%y")
```


## Problem 4

#### 4.a Write an R function, f4, that prints, with their row and column labels, only those elements of a correlation matrix for which $|\mbox{correlation}|\geq 0.9$.
```{r}
f4 <- function(df){
  # establish a null n*n matrix
  n <- ncol(df)
  n.names <- colnames(df)
  m.result <- matrix(nrow = n, ncol = n)

  # loop through all correlation permutations
  for(i in 1:n){
    for(j in 1:n){
      i.j.corr <- cor(df[i], df[j])
      m.result[i,j] <- i.j.corr
    }
  }
  rownames(m.result) <- n.names
  colnames(m.result) <- n.names
  return(m.result)
}
```

#### 4.b Using cars data, print the resuluts of f4.
```{r}
data(cars)
f4(cars)
```


## Problem 5

#### 5.a Derive the mean and variance of the Binomial distribution $(n,p)$.
$X = \sum\limits_{i=1}^n x_i$

$E(X) = E(\sum\limits_{i=1}^n x_i) = \sum\limits_{i=1}^n E(x_i) = \sum\limits_{i=1}^n 1*p(x_i=1) + 0*p(x_i=0) = \sum\limits_{i=1}^n p = np$

$Var(X) = \sum\limits_{i=1}^n [E(x_i^2) - E(x_i)^2] = \sum\limits_{i=1}^n [1^2p + 0^2(1-p) - (1*p + 0(1-p))^2] = \sum\limits_{i=1}^n p - p^2 = \sum\limits_{i=1}^n p(1-p) = np(1-p)$

#### 5.b Write an R function, f5, that simulates a student guessing at a True-False test consisting of 40 questions.
```{r}
f5 <- function(){
  options <- c(TRUE, FALSE)
  answers <- sample(options, 40, replace=TRUE)
  points <- 0
  log <- c()
  
  # simulate guessing
  for(i in 1:40){
    guess <- sample(options, 1)
    log <- c(log, guess)
    if(guess == answers[i]){
      points <- points + 1
    }
  }
  return(points)
}
```

#### 5.c Compute the number of student’s correct answers. Assuming $n=40$ and $p=0.5$, compare with the theoretical values.
```{r}
f5()
```
The expected value of Binomial(n=40, p=.5) = 20. We expect the student score 20 points. If there were more questions, the simulated result would more often be closer to the theoretical expected value.


## Problem 6. (optional)

##### 6.a
$P(x=k)=(1-p)^{k-1}p$

$F(x)=1-P(X>x)$

$F(x)=1-\sum\limits_{i=x+1}^{infinity}(1-p)^{s-1}p$

$F(x)=1-(1-p)^xp-(1-p)^{x+1}p-...$

$F(x)=1-(1-p)^xp[(1-p)^0+(1-p^1+(1-p)^2+...]$

$F(x)=1-(1-p)^xp[1/(1-(1-p))]$

$F(x)=1-(1-p)^xp/p=1-(1-p)^x=U$

$(1-p)^x=1-U$

$x=log_{1-p}(1-U)$

For p=.2, U=.62, $x=log_{.8}(.58)=2.44115$

##### 6.b
```{r}
u.sample <- runif(1000,0,1)
x.sample <- log(x=1-u.sample, base=1-.2)
hist(x.sample)
```
