---
output:
  pdf_document: default
  #html_document: default
---
### Nick Kurtansky

###  HW5: Permutation Test

**Due: Friday, 4/6/2018 (submit electronically in D2L Dropbox: No class on 4/6)**.

**Note:**  Students are required to submit their homework as both PDF file and hardcopy by the due date of the assignment.
Please turn in code separately and electronically. All electronic submissions should be made via D2L Dropbox and should follow the following naming convention: last name, first name, assignment number, proper extension. So, for example, if John Smith is turning in Homework 1, he would name the file Smith_John1.pdf. The associated code would be Smith_John1.Rmd. If you wish to break up your code into separate files, you may submit them as Smith_John1a.Rmd, Smith_John1b.Rmd, and so on. 
There will be a 20%  penalty per day that your homework is late.  Homework in the wrong format will not be given credit.

```{r include=FALSE}
library(dplyr)
```

**Problem 1**

Suppose we are interested in the effects of *fertilization*  on the *species richness* of invertebrates in a lake. 

Lake|Richness|Area|Fertilized
---|---|---|------------|
1|32|2.0|1
2|29|0.9|1
3|35|3.1|1
4|36|3.0|1
5|41|1.0|0
6|62|2.0|0
7|88|4.0|0
8|77|3.5|0
mean|50|2.4375|0.5
SD|22.6|1.1426|0.535


1)  Create a data frame using `data.frame` for a given data. Define the variable `Rich2` which is the binary indicator if  `Richness`$>$ 40 then 1, 0 otherwise. 
To answer if there is an effect of fertilization on `Rich2`, we can set up a null hypothesis as follows:
```{r}
Lake <- seq(1,8,1)
Richness <- c(32, 29, 35, 36, 41, 62, 88, 77)
Area <- c(2.0, 0.9, 3.1, 3.0, 1.0, 2.0, 4.0, 3.5)
Fertilized <- c(1, 1, 1, 1, 0, 0, 0, 0)
Rich2 <- rep(NA, 8)

# Initialize data frame and create new variable Rich2
data <- data.frame(Lake, Richness, Area, Fertilized, Rich2)
for(i in 1:nrow(data)){
  if(data$Richness[i] > 40){
    data$Rich2[i] = 1
  } else{
    data$Rich2[i] = 0
  }
}
data
```

$H_0: \beta_1=0,$ where $\beta_1$ is the slope coefficient in the logistic regression model. 

$H_a: \beta_1\neq0$

Alpha = .05

2)  Obtain $\beta_1$ by fitting logistic regression. 
```{r}
# logistic regression in R
fit <- glm(data$Rich2 ~ data$Fertilized, family=binomial(link=logit))
summary(fit)
```
 $\beta_1 = -49.13$
 
3) Obtain the empirical $p$-value.
```{r warning=FALSE}
set.seed(10)

B <- 10000
nu <- 1:8
reps <- numeric(B)
#absolute value of correlation because we are conducting a two-sided test (cor could be - or +)
rho.0 <- (cor.test(data$Fertilized, data$Rich2, method = "spearman")$estimate)

  #begin permutations test
for(i in 1:B){
  perm <- sample(nu, size=8, replace = FALSE)
  fert.shuffle <- data$Fertilized[perm]
  
  reps[i] <- (cor.test(data$Rich2, fert.shuffle, method = "spearman")$estimate)
}

p.1 <- 2 * mean(c(rho.0, reps) <= rho.0)  #calculate p-value from the permutations
p.1
```

4) Do we reject the null hypothesis?

The observed p-value is .0282 is less than .05, therefore we reject the null hypothesis that there is no association between fertilization and lake richness.

**Problem 2**

1) Implement the bivariate Spearman rank correlation test for independence as a permutation test.
You can generate correlated samples using the `mvrnorm` function. For example,

Ho: The columns are NOT correlated ($\rho = 0$)

Ha: The columns are associated with one antoher in some sort of way ($\rho \neq 0$)
alpha = .05

```{r}
set.seed(14)

#code given by professor to generate sample
library(MASS)
mu<-c(0,0)
Sigma<- matrix(c(1,0.5,0.5,1),2,2)
n<-30
x<-mvrnorm(n,mu,Sigma)

  #establish each column as its own vector
col.1 <- x[,1]
col.2 <- x[,2]

#begin permutation test set-up
B=10000
  #sampling from 30 rows
id <- 1:30
R <- numeric(B)
r.0 <- (cor.test(col.1, col.2, method="spearman")$estimate)
#begin purmutation iteration!
for(i in 1:B){
  perm <- sample(id, size = 30, replace = FALSE)
  col.2.star <- col.2[perm]
  R[i] <- (cor.test(col.1, col.2.star, method = "spearman")$estimate)
}
  #p-value from permutation test
p.2 <- 2 * mean(c(r.0, R)>=r.0)
p.2
```

2)
Compare the achieved significance level of the permutation test with the
$p$- value reported by `cor.test (method="spearman")` on the same samples.
```{r}
  #p-value from Spearman Correlatino Test
cor.test(col.1, col.2, method = "spearman")$p.value

  #p-value from Permutations Test
p.2
```

These samples are moderately correlated with one-another with rho = .59555. The permutation test produced a p-value of .0012, which is slightly larger than the Spearman correlation test's p-value of .00657. The p-value is significantly less than .05 in both cases. Therefore we conclude that the samples are not independent from one-another; there is some correlation.



